r5r: Rapid Realistic Routing with R5 in R

### 2020-09-17 r-package/paper2020/paper2020.Rmd
updates travel time matrix paper section

- An OSM network in `.pbf` format (*mandatory*).
- A public transport network in `GTFS.zip` format.
- The spatial coordinates of points covering the area in `.csv` format.

The `.pbf` and `GTFS.zip` files are used to build the transport network. While 
the first is mandatory, since it is required to create a street network, the 
latter is only used to calculate public transport trips, thus may not be 
necessary if that's not intended.

@@

### Build routable transport network

The first step is to build the multimodal transport used for routing in R5. This
is done with the `setup_r5` function. It does two things: (1) downloads a compiled JAR file
of R5 and store it locally in the r5r package directory; (2) combines the osm.pbf and gtfs.zip data sets to build a routable network object. The function

```{r, message = FALSE}
# Indicate the path where OSM and GTFS data are stored
r5r_core <- setup_r5(data_path)
```

**r5** has two core routing functions.

- `detailed_itineraries`
   * Returns a `data.frame sf LINESTRINGs` with one or multiple alternative routes
   between one or multiple origin destination pairs. The data output brings 
   detailed information on transport mode, travel time, walk distance etc for 
   each trip section

- `travel_time_matrix`
   * Fast function that returns a simple 'data.frame' with travel time 
   estimates between one or multiple origin destination pairs.

## Fast many to many travel time matrix

```{r, message = FALSE}
mode <- c("WALK", "TRANSIT")
max_walk_dist <- Inf
max_trip_duration <- 120L
departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S")

# estimate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode,
                          departure_datetime,
                          max_walk_dist,
                          max_trip_duration)

head(ttm)
```

## Detailed itineraries

>>>>

- An OSM network in `.pbf` format.
- A public transport network in `GTFS.zip` format.
- The spatial coordinates of points covering the area in `.csv` format.

@@

### Build a routable transport network

Building a transport network with r5r is as simple as a one command line:

```{r, message = FALSE}
r5r_core <- setup_r5(data_path)
```

The function uses the `.pbf` and the `GTFS.zip` files in the directory pointed 
by `data_path` to create a multimodal transport network used for routing by 
R<sup>5</sup>. The resulting `network.dat`, as well as some other files used by 
R<sup>5</sup>, is saved inside the supplied directory.

While the presence of a `.pbf` file is mandatory, since it is required to create 
a street network, the `GTFS.zip` file is only used to calculate public transport 
trips, thus may not be necessary if that's not intended. Multiple GTFS can be 
used simultaneously, in which case all files are merged into a single one 
automatically.

### Calculate a travel time matrix

Calculating a travel time matrix with r5r can be done very efficiently with the
`travel_time_matrix()` function. The function takes travel restrictions and
origins/destinations spatial location as inputs and outputs a `data.table` with 
the travel time between all combinations of origins and destinations that can be 
reached with the given restrictions:

```{r, message = FALSE}
# set restrictions
mode <- c("WALK", "TRANSIT")
max_walk_dist <- 1000
max_trip_duration <- 120
departure_datetime <- as.POSIXct("13-03-2019 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S")

# calculate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode,
                          departure_datetime,
                          max_walk_dist,
                          max_trip_duration)

ttm
```

Note that the resulting `ttm` object only holds the travel time of 1043 
origin/destination pairs, even though the `points` data set passed both to 
`origins` and `destinations` argument contains 100 points. This is because out 
of the total 10000 possible trips between the supplied points, only 1043 could 
be completed within the specified restrictions. Removing any walking distance 
limits (by setting it to infinite) and increasing maximum trip duration 
accordingly results in a "complete" output?

```{r, message = FALSE}
# set softer restrictions
max_walk_dist <- Inf
max_trip_duration <- 360

# calculate travel time matrix
ttm <- travel_time_matrix(r5r_core,
                          origins = points,
                          destinations = points,
                          mode,
                          departure_datetime,
                          max_walk_dist,
                          max_trip_duration)

ttm
```

### Detailed itineraries
